Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-19 doc/GMP2015/s04_recombination.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

In this section, we consider the \dfn{optimal recombination} problem
of combining specific solutions of subsystems in a DR-plan into a
solution of their parent system $I$, (without loss of generality, at
the top level of the DR-plan). In the case of isostatic qusecs, the
parent system $I$ is isostatic (the root of the DR-plan), and we seek
\dfn{solution(s) (among a finite large number of solutions) with a
specific orientation or chirality}. In the case of underconstrained
qusecs the subsystems are the multiple roots of the DR-plan, the
parent system $I$ is underconstrained, and we typically seek an
efficient algorithmic description of \dfn{connected component(s) of
solutions with a specific orientation or chirality}.

The main barrier in recombination when given an optimal DR-plan (of
smallest possible size or max fan-in),  is that the number of children
(resp. number of roots) - and correspondingly the  size and complexity
of the (indecomposable) algebraic system $I$ to be solved - could be
arbitrarily large as a function of the size of the input system. This
difficulty can persist even after optimal parametrization of the
indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic
complexity.

In the following, we formulate the problem of \dfn{optimal
modification}, of the indecomposable algebraic system $I$ into a
decomposable system with a small DR-plan (low algebraic complexity).
Leveraging recent results on \dfn{Cayley configuration spaces}, our
approach to the optimal modification problem achieves the following:
\dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters
$F$,  all of which have small DR-plans. Thus, given a value $v$ for
$\lambda_F$, the system $I_v$ can potentially be solved or realized
easily once the orientation type of the solution is known  (when the
DR-plan size is small enough). \dfn{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions. \dfn{(c) Efficient search.} Finally, this union can be projected into the
so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
convex or otherwise easy to traverse in order to search for $I$'s
solution (or connected component) of the desired orientation type. For
the case when the modification is bounded, this approach provides an
efficient algorithm for recombination. We first define the decision
version of the problem of optimal modification for decomposition. The
standard optimization versions are straightforward.

\problemstatement{Optimal Modification for Decomposition (OMD) Problem.}
Given a generically independent graph $G = (V,E)$ with no non-trivial
proper isostatic subgraph (indecomposable), and 2 constants $k$ and
$s$, does there exist a set of at most $k$  edges $E_1$ and a set of
nonedges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD
where $k$ is a fixed bound (not part of the input).   We say that such
a tuple $(G,s)$ is a member of the set OMD$_k$. We loosely refer to
graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
with appropriately small $s$.

It is immediately clear that indecomposable graphs $G$ that belong in
OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
decomposable graphs satisfying Criteria (a) and (b) above. However, it
is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
\label{sec:2-tree-reduction}
%
Next we provide the necessary background to describe a specific
approach for achieving the requirements (a)-(c) mentioned above, by
restricting the class of reduced graphs $G' = G\setminus E_1$ and
their isostatic completions $H$ in the above definition of the OMD
problem, and using a key theorem of Convex Cayley configuration spaces
\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-
edges $F$ (Cayley parameters), such that the set of vectors
$\lambda_F$ of  attainable lengths of the nonedges $F$ is always
convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
all the realization s of the bar-joint constraint system or linkage
$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
a ``projection'' of the cartesian realization space of $(H,\delta)$ on
the Cayley parameters $F$. Such graphs $H$ are said to have
\dfn{convexifiable Cayley configuration spaces with parameters $F$}
(short: \dfn{convexifiable}). To state the theorem, we first have to
define the notion of \dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and
$H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with
edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
respectively.  A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum
components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one
that cannot be further split into 2-sum components. A \dfn{2-tree} is
recursively obtained by taking a 2-sum of two 2-trees, with the base
case of 2-tree being a triangle. A \dfn{partial 2-tree} is a 2-tree
minus some edges. Partial 2-trees are characterized as having $K_4$ as
forbidden minor are also called series parallel graphs.

@@

The idea of our approach to achieve the criteria (a)-(c) begins with
the following simple but useful theorem.

@@

We refer to such graphs $G$ in short as \dfn{$k$-approximately
convexifiable}, where the reduced graphs $G'$ and isostatic
completions $H$ are convexifiable. As observed earlier, since graphs
such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met
for small enough $k$. Criterion $c$ is addressed as described in the
following efficient search procedure which clarifies the dependence of
the complexity on the number and ranges  of Cayley parameters $F$.

\begin{theorem}\label{theorem:criterionc}
    [Efficient search]
    For an indecomposable, $k$-approximately convexifiable graph $G =
    (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial
    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of nonedges
    of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution
    $p$ (or connected component of a solution space) of $(G,\delta)$
    of an orientation type $\sigma_p$ can be found in time
    $O(\log(W))$ where $W$ is the number of cells of desired accuracy
    (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$.
    The (discrete) volume $W$ is exponential in $|F|$ and polynomial
    in the (discrete range) of the parameters in $F$.
\end{theorem}

@@

\myexample
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the nonedge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

>>>>

In this section, we consider the \dfn{optimal recombination} problem
of combining specific solutions of subsystems in a DR-plan into a
solution of their parent system $I$ (without loss of generality, at
the top level of the DR-plan). In the case of isostatic qusecs, the
parent system $I$ is isostatic (the root of the DR-plan), and we seek
\dfn{solution(s) (among a finite large number of solutions) with a
specific orientation or chirality}. In the case of under-constrained
qusecs the subsystems are the multiple roots of the DR-plan, the
parent system $I$ is under-constrained, and we typically seek an
efficient algorithmic description of \dfn{connected component(s) of
solutions with a specific orientation or chirality}.

The main barrier in recombination when given an optimal DR-plan (of
smallest possible size or max fan-in),  is that the number of children
(resp. number of roots) --- and correspondingly the  size and complexity
of the (indecomposable) algebraic system $I$ to be solved --- could be
arbitrarily large as a function of the size of the input system. This
difficulty can persist even after optimal parametrization of the
indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic
complexity.

In the following, we formulate the problem of \dfn{optimal
modification}, of the indecomposable algebraic system $I$ into a
decomposable system with a small DR-plan (low algebraic complexity).
Leveraging recent results on \dfn{Cayley configuration spaces}, our
approach to the optimal modification problem achieves the following:
\dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ ---  one for each value $\lambda_F$ for the parameters
$F$,  all of which have small DR-plans. Thus, given a value $v$ for
$\lambda_F$, the system $I_v$ can potentially be solved or realized
easily once the orientation type of the solution is known  (when the
DR-plan size is small enough). \dfn{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions. \dfn{(c) Efficient search.} Finally, this union can be projected into the
so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
convex or otherwise easy to traverse in order to search for $I$'s
solution (or connected component) of the desired orientation type. For
the case when the modification is bounded, this approach provides an
efficient algorithm for recombination. We first define the decision
version of the problem of optimal modification for decomposition. The
standard optimization versions are straightforward.

\problemstatement{Optimal Modification for Decomposition (OMD) Problem.}
Given a generically independent graph $G = (V,E)$ with no non-trivial
proper isostatic subgraph (indecomposable) and 2 constants $k$ and
$s$, does there exist a set of at most $k$  edges $E_1$ and a set of
non-edges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD
where $k$ is a fixed bound (not part of the input).   We say that such
a tuple $(G,s)$ is a member of the set OMD$_k$. We loosely refer to
graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
with appropriately small $s$.

It is immediately clear that indecomposable graphs $G$ that belong in
OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
decomposable graphs satisfying Criteria (a) and (b) above. However, it
is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
\label{sec:2-tree-reduction}
%
Next we provide the necessary background to describe a specific
approach for achieving the requirements (a)--(c) mentioned above, by
restricting the class of reduced graphs $G' = G\setminus E_1$ and
their isostatic completions $H$ in the above definition of the OMD
problem, and using a key theorem of Convex Cayley configuration spaces
\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$
and non-edges $F$ (Cayley parameters), such that the set of vectors
$\lambda_F$ of  attainable lengths of the non-edges $F$ is always
convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
all the realizations of the bar-joint constraint system or linkage
$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
a ``projection'' of the Cartesian realization space of $(H,\delta)$ on
the Cayley parameters $F$. Such graphs $H$ are said to have
\dfn{convexifiable Cayley configuration spaces with parameters $F$}
(short: \dfn{convexifiable}).

To state the theorem, we first have to
define the notion of \dfn{2-sums} and \dfn{2-trees}.
Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.
% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum
components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one
that cannot be further split into 2-sum components. A \dfn{2-tree} is
recursively obtained by taking a 2-sum of 2-trees, with the base
case being a triangle. A \dfn{partial 2-tree} is a 2-tree
minus some edges. Partial 2-trees characterized as having $K_4$ as a
forbidden minor are also called series parallel graphs.

@@

The idea of our approach to achieve the criteria (a)--(c) begins with
the following simple but useful theorem.

@@

We refer to such graphs $G$ in short as \dfn{$k$-approximately
convexifiable}, where the reduced graphs $G'$ and isostatic
completions $H$ are convexifiable. As observed earlier, since graphs
such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met
for small enough $k$. Criterion (c) is addressed as described in the
following efficient search procedure which clarifies the dependence of
the complexity on the number and ranges  of Cayley parameters $F$.

\begin{theorem}\label{theorem:criterionc}
    [Efficient search]
    For an indecomposable, $k$-approximately convexifiable graph $G =
    (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial
    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of non-edges
    of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution
    $p$ (or connected component of a solution space) of $(G,\delta)$
    of an orientation type $\sigma_p$ can be found in time
    $O(\log(W))$ where $W$ is the number of cells of desired accuracy
    (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$.
    The (discrete) volume $W$ is exponential in $|F|$ and polynomial
    in the (discrete range) of the parameters in $F$.
\end{theorem}

@@

\myexample
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
non-edge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the non-edge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met --- for example when $k=1$ --- because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

### 2014-12-19 doc/GMP2015/s05_problems.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

\section{Problem Relationships and Unification}
\label{sec:table}
In this section we provide a unified view of  the various problems
studied in the previous 2 sections, along with formal reductions
between them, and relationships to other known problems and results as
well as open questions.
%
\subsection{Special classes of small DR-plans}
As seen in the previous section, 2-trees and tree-decomposable graphs
have not only small, but also special DR-plans that permit easy
solving - essentially a single quadratic at a time.

@@

%In the recombination setting, the corresponding \dfn{restricted
%OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic
%completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction}
%or to be a low Cayley complexity tree-decomposable graph as in Section
%\ref{sec:tdecomp}. Clearly these problems have deterministic
%polynomial time algorithms, the algorithms run in time exponential in
%$k$. As a result, when $k$ is part of the input, the \dfn{restricted
%OMD problem} has the potential to be difficult. For example, when the
%isostatic completion $H$ is required to be a 2-tree the restricted OMD
%problem is reducible to the maximum spanning series-parallel subgraph
%problem shown by \cite{cai1993spanning} to be NP-complete even if the input
%graph is planar of maximum degree at most 6. However, since the OMD
%problem has other input restrictions such as not having any proper
%isostatic subgraphs, it is not clear if the reverse reduction exists
%and hence it is unclear whether the OMD problem is NP-complete. The
%same holds for the restricted OMD problem where the isostatic
%completion $H$ is required to be a tree-decomposable graph of low
%Cayley complexity (i.e, have special, small DR-plans). One potential
%obstacle to an indecomposable graph $G$'s membership in the restricted
%OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
%In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
%have been constructed in \cite{servatius2000rigidity}.
%
\subsection{Optimal Modification, Completion and Recombination: formal
connections}
%
The OMD problem is closely related to a well-studied problem of
completion of an under-constrained system to an isostatic one with a
small DR-plan.
\begin{observation}\label{obs:OC_to_OMD}
    The (decision version of) the {\bf optimal completion problem
    (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
\end{observation}
In fact, a \dfn{restricted OC} problem was studied by \cite{joan-arinyo2003transforming}
requiring the completion to be tree-decomposable.
%It is not clear if
%the reverse reduction holds, i.e, is OMD reducible to OC?

@@

In order to connect the OR problem to OMD, when the input graph for is
the isostatic graph at the DR-plan root, we do not consider the case
where the two child \dfn{solved subgraphs} (corresponding to already
solved subsystems) have a nontrivial intersection (in this case the
recombination is trivial). We only consider the case where no two
child solved subgraphs (resp. two root subgraphs when the input graph
is under-constrained) share more than 1 vertex. We replace such solved
subgraphs  by isostatic graphs as follows. If a solved subgraph shares
at most one vertex with the remainder of the graph, simply replace it
by an edge one of whose endpoints  is the shared vertex. Otherwise,
replace it by  a 2-tree graph of the shared vertices. Finally, we add
the additional restriction to the OM Problem that when any edge in a
solved subgraph is chosen among the $k$ edges to be removed, in fact
the entire solved subgraph must be removed  and all of its edges must
be counted in $k$.

This reduction is used also for transferring  algorithms for optimal
DR-planning, recombination, completion, OMD, and other problems
from bar-joint systems to
so-called \dfn{multi-body pin} define in Section \ref{sec:bodypin}
by showing that the corresponding problems for the latter are reduced
to the corresponding problems on bar-joint systems.

>>>>

\section{Problem Relationships and Unification}
\label{sec:table}
In this section we provide a unified view of  the various problems
studied in the previous 2 sections, along with formal reductions
between them, and relationships to other known problems and results as
well as open questions.
%
\subsection{Special classes of small DR-plans}
As seen in the previous section, 2-trees and tree-decomposable graphs
have not only small, but also special DR-plans that permit easy
solving --- essentially a single quadratic at a time.

@@

%In the recombination setting, the corresponding \dfn{restricted
%OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic
%completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction}
%or to be a low Cayley complexity tree-decomposable graph as in Section
%\ref{sec:tdecomp}. Clearly these problems have deterministic
%polynomial time algorithms, the algorithms run in time exponential in
%$k$. As a result, when $k$ is part of the input, the \dfn{restricted
%OMD problem} has the potential to be difficult. For example, when the
%isostatic completion $H$ is required to be a 2-tree the restricted OMD
%problem is reducible to the maximum spanning series-parallel subgraph
%problem shown by \cite{cai1993spanning} to be NP-complete even if the input
%graph is planar of maximum degree at most 6. However, since the OMD
%problem has other input restrictions such as not having any proper
%isostatic subgraphs, it is not clear if the reverse reduction exists
%and hence it is unclear whether the OMD problem is NP-complete. The
%same holds for the restricted OMD problem where the isostatic
%completion $H$ is required to be a tree-decomposable graph of low
%Cayley complexity (i.e, have special, small DR-plans). One potential
%obstacle to an indecomposable graph $G$'s membership in the restricted
%OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
%In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
%have been constructed in \cite{servatius2000rigidity}.
%
\subsection{Optimal Modification, Completion and Recombination: formal
connections}
%
The OMD problem is closely related to a well-studied problem of
completion of an under-constrained system to an isostatic one with a
small DR-plan.
\begin{observation}\label{obs:OC_to_OMD}
    The (decision version of) the \textbf{optimal completion problem
    (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
\end{observation}
In fact, a \dfn{restricted OC} problem was studied by \cite{joan-arinyo2003transforming}
requiring the completion to be tree-decomposable.
%It is not clear if
%the reverse reduction holds, i.e, is OMD reducible to OC?

@@

In order to connect the OR problem to OMD, when the input graph is
the isostatic graph at the DR-plan root, we do not consider the case
where the two child \dfn{solved subgraphs} (corresponding to already
solved subsystems) have a nontrivial intersection (in this case the
recombination is trivial). We only consider the case where no two
child solved subgraphs (resp. two root subgraphs when the input graph
is under-constrained) share more than 1 vertex. We replace such solved
subgraphs  by isostatic graphs as follows. If a solved subgraph shares
at most one vertex with the remainder of the graph, simply replace it
by an edge one of whose endpoints  is the shared vertex. Otherwise,
replace it by  a 2-tree graph of the shared vertices. Finally, we add
the additional restriction to the OM Problem that when any edge in a
solved subgraph is chosen among the $k$ edges to be removed, in fact
the entire solved subgraph must be removed  and all of its edges must
be counted in $k$.

This reduction is used also for transferring  algorithms for optimal
DR-planning, recombination, completion, OMD, and other problems
from bar-joint systems to
so-called \dfn{body-hyperpin}, defined in Section \ref{sec:bodypin},
by showing that the problems for the latter are reduced
to the corresponding problems on bar-joint systems.

### 2014-12-19 doc/GMP2015/s07_problempinning.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

We can use qusecs DR-plans to design materials such as disordered
graphene and silica bi-layers \cite{silica_bilayers}
\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
a somewhat more general setting:
the problem of finding boundary conditions (additional constraints) to
add to an underconstrained monolayer to make it
isostatic.
This can be done in a number of ways.

@@

\begin{enumerate}
    \item Pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
    \item Pin the boundary of (or in general, add constraints to) a layer
        (possibly a genus 0 monolayer) so that it becomes isostatic
    \item Design a broader class of structures to ensure they are isostatic,
        self-similar (via some subdivision rule) and in addition isostatic at each
        level of the subdivision (see Figure \ref{fig:subdivision})
\end{enumerate}

@@

\begin{definition}
\label{def:body-pin}
    A body-pin graph is a body-hyperpin graph with the following conditions:
    \begin{itemize}
        \item Each pin is shared by at most 2 bodies
        \item No 2 bodies share more than one pin
    \end{itemize}
    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where
    the bodies are the original bodies and each pin represents 2 bars from
    one body to another. Such body-bar graphs with 1 and 2-dof can be
    characterized by being $(3,4)$ and $(3,5)$-tight respectively
    \cite{Lee:2007:PGA} \cite{streinu2009sparse},
defined in Section  \ref{sec:preliminaries.}
\end{definition}

\begin{definition}
    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body
    is a triangle, i.e., it shares at most 3pins with other bodies. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hypergraphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

@@

We can develop the notion of a canonical DR-plan when the hypergraph is
$(k,l)$-sparse (i.e., combinatorially independent) using certain
notion of trivial and non-trivial intersections and
$(k,l)$-tightness conditions similar to the way we did in Section
\ref{sec:DRP}, provided
    such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
    In particular, we can define canonical DR-plans with similar properties
    for the 1 and 2 dof body-pin and triangle-hyperpin systems defined
    above.

\begin{observation}
\label{obs:bodypin_drp}
    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity
    corresponds to canonical DR-plan node
    satisfies one of the following:

    \begin{itemize}
        \item Its children  are 2 proper vertex maximal 1-dof graphs that intersect on
            another 1-dof graph
        \item Its children are a number of proper maximal 1-dof sub-graphs,
            sharing at most one body
    \end{itemize}
\end{observation}

@@

\begin{observation}
\label{obs:algebraic_completion}
    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for
    the isostatic completion with minimum fanin, (a reasonable
    measure of algebraic complexity), a more nuanced
    measure that treats solutions of 1-dof and 2-dof systems as
    1- or 2- parameter families would no longer permit the algorithm given
    in the proof. In particular the complexity of the standard algorithm in the k-dof
    case would be exponential in $k$ (even if that case were matroidal and
    an optimal DR-plan is known).
\end{observation}

>>>>

@@

We can use qusecs DR-plans to design materials such as disordered
graphene and silica bi-layers \cite{silica_bilayers}
\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
a somewhat more general setting:
the problem of finding boundary conditions (additional constraints) to
add to an underconstrained monolayer to make it
isostatic.
This can be done in a number of ways:
(1) pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass});
(2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or
(3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).

@@

\begin{definition}
\label{def:body-pin}
    A body-pin graph is a body-hyperpin graph with the following conditions:
    (1) each pin is shared by at most two bodies; and
    (2) no two bodies share more than one pin

    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where
    the bodies are the original bodies and each pin represents 2 bars from
    one body to another. Such body-bar graphs with 1 and 2-dof can be
    characterized by being $(3,4)$ and $(3,5)$-tight respectively
    \cite{Lee:2007:PGA} \cite{streinu2009sparse},
defined in Section  \ref{sec:prelim}.
\end{definition}

\begin{definition}
    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body
    is a triangle, i.e., it shares at most 3 pins with other bodies. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hypergraphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

@@

We can develop the notion of a canonical DR-plan when the hypergraph is
$(k,l)$-sparse (i.e., combinatorially independent) using certain
notion of trivial and non-trivial intersections and
$(k,l)$-tightness conditions similar to the way we did in Section
\ref{sec:DRP}, provided
    such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
    In particular, we can define canonical DR-plans with similar properties
    for the 1 and 2-dof body-pin and triangle-hyperpin systems defined
    above.

\begin{observation}
\label{obs:bodypin_drp}
    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity corresponds to a canonical DR-plan whose nodes satisfy one of the following, either
    (a) the children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph, or
    (b) the children are a number of proper maximal 1-dof sub-graphs, sharing at most one body.
\end{observation}

@@

\begin{observation}
\label{obs:algebraic_completion}
    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for
    the isostatic completion with minimum fan-in, (a reasonable
    measure of algebraic complexity), a more nuanced
    measure that treats solutions of 1-dof and 2-dof systems as
    1- or 2- parameter families would no longer permit the algorithm given
    in the proof. In particular the complexity of the standard algorithm in the k-dof
    case would be exponential in $k$ (even if that case were matroidal and
    an optimal DR-plan is known).
\end{observation}

### 2014-12-19 doc/GMP2015/s09_conclusion.tex
Fixed some references and citations. Unified display of OptDRP, ComDRP, CanDRP, use \drp{G}, \optdrp{G}, etc.
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

\section{Conclusion}
We have worked to navigate the complexity barriers faced by the general DR-plan and recombination problem. To this end we have defined a new canonical DR-plan and given an effiecient ($O(n^2)$) algorithm to find the DR-plan for specific types of 2D qusecs.

We have also described methods of recombining the DR-plan to get a realization different from the normal solving of a system of equations. These results rely on the results obtained in the Cayley Configuration space of certain graphs.

We then related these theoretical results to real-world situations in which they can be applied including Examples 1-5 in Section \ref{sec:intro}.

Throughout the paper, we raised some open questions and future areas of research. These include: completion of body-hyper-pin qusecs with more than 2-dof, \todo{Add a general synopsis of other questions}

>>>>

\section{Conclusion}
\label{sec:conclusion}
% \bibliography{paper}

We have elucidated the complexity barriers, for example over-constraints, faced by the general optimal DR-plan and recombination problem. When there are no overconstraints, which is the case for 2D qusecs whose realizations are many common types of layered materials, we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property given an efficient ($O(n^2)$) algorithm to find a DR-plan that satisfies other desirable characteristics.

We have also described novel methods of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise non-decomposable system that recombines the solved child systems at a node of a DR-plan. These results rely on a recent theory of Convex Cayley Configuration spaces.

We then modeled specific layered materials using extensions of the above theoretical results including the motivating Examples 1-5 in the introduction.

In Appendix B [\ref{sec:appendix:b}] we provide open problems for future work.
